---
title: "manuscript data and figures"
author: "Katie Palof"
date: "September 12, 2017"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=F}
library(knitr)
library(tidyverse)
library(FSA)
library(gridExtra)
library(weights)
library(nlstools)
library(AICcmodavg)
library(ggplot2)
options(scipen=9999) # remove scientific notation

knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)
```
Housekeeping/ long-term: code is in geoduck_exploration R project. See regional_age_growth_cc.R and regional_catchcurve.R

## Area specific trends
```{r data load, echo = FALSE, message = FALSE, warning=FALSE}
## Load data -------------------------------------------------------
dat <- read.csv("../data/12_14_geoduck_all.csv")
weight_pop <- read.csv("../data/GeoduckAgeStudyWeighting.csv")

weight_pop %>%  filter(area != "Biorka/Legma Islands") %>% 
  filter(area != "Taigud/Kolosh Islands") -> weight_pop #remove sitka area since they are grouped under "sitka" now
weight_pop %>% mutate(wt_each = popsize_wshow/(sum(popsize_wshow))) %>% 
  select(-AgeStudyWeighting)-> weight_pop

# need to edit data to make sitka one sample - so group Biorka/Legma Islands and Taigud/Kolosh Islands together
#unique(dat$ADFG_Fishery.Area)
dat %>% mutate(area1 = as.character(ADFG_Fishery.Area))  %>% 
        mutate(area = ifelse(area1 == "Biorka/Legma Islands", "Sitka", 
                             ifelse(area1 == "Taigud/Kolosh Islands", 
                                    "Sitka", area1))) %>% 
        select( -area1) -> dat1
  
# full data set with weigtings
dat1 %>% left_join(weight_pop) -> dat2 # raw data
# summarized by area, age with weigthing
dat2 %>% group_by(area, Age_2012) %>% 
  summarise(n = n())  -> dat2_by.area # counts by age for each area 
dat2_by.area %>% left_join(weight_pop) %>% mutate(n_wt = n*wt_each) -> dat_wt_by.area # counts by age/area with weighting

## weighting counts ------------------
dat1 %>% group_by(area) %>% summarise(N_samp = n()) ->total.n.area  # number sampled in each area
dat_wt_by.area %>% left_join(total.n.area) %>% 
  mutate(prop = n/N_samp, n_corrected = prop*popsize_wshow) -> dat_wt_by.area2
# n_corrected is counts using population size weighting.  n_wt is proportion using same weighting
# this is just a scale peference

################ data sets by sea otter ---------------
dat2 %>% filter(otter.status == "present") -> present_raw
dat2 %>% filter(otter.status == "absent") -> absent_raw
dat_wt_by.area2 %>% filter(otter.status == "present") -> present_wt
dat_wt_by.area2 %>% filter(otter.status == "absent") -> absent_wt
```

### Unweighted histograms
```{r area histograms }
ggplot(dat2, aes(Age_2012, fill = otter.status)) + geom_histogram(binwidth = 1.0, alpha = 0.5, position = "dodge") +
  facet_grid( area ~.)
```

### Weighted mean age 
```{r area means }
# use dat_wt_by.area2  weight mean age by n_corrected or n_wt(weight)
dat_wt_by.area2 %>% group_by(area) %>% summarise(mean = weighted.mean(Age_2012, n_wt), 
               SE = (wt.sd(Age_2012, n_wt)/(sqrt(sum(!is.na(Age_2012))))), 
               MIN = min(Age_2012, na.rm = TRUE), 
               MAX = max(Age_2012, na.rm = TRUE), 
               n = sum(n))
```


### Means, histograms (weighted or unweighted???), growth?




## Regional groupings
Geoduck age collections were aggregated into regional groupings based on geography, environment, and present of predators (specifically the sea otter).  Group 1 includes the Nakat Inlet, Vallenar Bay, and Vegas/Hotspur Islands, which are all absent of sea otters (total sample size = 611).  Group 2 includes Cone Island North, East San Fernando, Tlevak Strait, Warren/Kosciusko Islands, and Sitka (Biorka and Taigud), which all have sea otters present (total sample size = 1048).  For purposes of this document Group 1 will be referred to as the "absent" group and Group 2 will be "present".  (we can change this later if needed - this is just a draft)

### Weighted mean age 
```{r mean age by group}
# use dat_wt_by.area2  weight mean age by n_corrected or n_wt(weight)
dat_wt_by.area2 %>% group_by(otter.status) %>% summarise(mean = weighted.mean(Age_2012, n_wt),
                       SE = (wt.sd(Age_2012, n_wt)/(sqrt(sum(!is.na(Age_2012))))), 
                       MIN = min(Age_2012, na.rm = TRUE), 
                       MAX = max(Age_2012, na.rm = TRUE), 
                       n = sum(n))
```

## Age Composition Similar?
More exploration is neeeded for this analysis but a simple ks test suggests they are not from the same distribution. I'm currently researching other tests/ analyses to look at this, and i also think I should bin the ages and redo this test just for sensitivity.  (Code ee is present and ff is absent.)

```{r}
present_dist <- data.frame(x = rep(present_wt$Age_2012, present_wt$n_corrected), z = rep(present_wt$otter.status, present_wt$n_corrected))

absent_dist <- data.frame(x = rep(absent_wt$Age_2012, absent_wt$n_corrected), z = rep(absent_wt$otter.status, absent_wt$n_corrected))

ks.test(present_dist$x, absent_dist$x) # reject null that two are equal

present_dist %>% bind_rows(absent_dist) -> all_n_corrected # expanded counts combined into one data set. x is ages and z is otter.status
```

## All age samples combined
Geoduck age samples from 2012 and 2014 were combined into one dataset to estimate natural mortality, etc. 